An Analysis of the Feasibility and Performance of BitTorrent as a Data Transfer Protocol in Peer to Peer Backup Systems

### 2014-04-23 paper/paper.tex
Paper: analyze results of test scenario 1

\begin{figure}
  \centering
  \includegraphics[scale=0.75]{PeerStoreTrading}
  \caption{Trading protocol of PeerStore \cite{PeerStore} \label{fig:peerStoreTrading}}
\end{figure}

@@

\begin{figure} \label{fig:convergent}
  \centering
  \includegraphics[scale=0.75]{ConvergentEncryption}
  \caption{Convergent Encryption from PeerStore}
\end{figure}

@@

\section{The BitTorrent Protocol} \label{sec:TheBitTorrentProtocol}
BitTorrent is a peer to peer protocol designed for fast file sharing. A peer who wishes to share a file creates a .torrent file containing metadata about the file to be shared. This peer is known as the file's initial \textit{seeder}, or uploader, and through the use of a BitTorrent client, starts \textit{seeding}, or uploading, the file. When a peer acts as a seeder, its sole purpose is to share the file that it is seeding with other peers. Peers who wish to download the file must acquire its .torrent file through an out-of-band channel, which is often a torrent indexer. Peers use the .torrent file to locate the initial seeder through either a centralized tracker or a DHT. These peers then connect to the initial seeder and start downloading the file. These peers are known as \textit{leechers} until they possess a full copy of the file, which is when they, themselves, become seeders.

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.4]{SystemDesign}
  \caption{The design of our system \label{fig:SystemDesign}}
\end{figure}

@@

In order to thoroughly evaluate our system, we first identified several variables that could lead to differences in performance. These variables are:

@@

From these variables, we derived a set of test scenarios which allowed us to see how each of them affect the performance of the system. While they are specific to our system, they allow us to evaluate the strengths and weaknesses of BitTorrent Sync (and, by extension, BitTorrent) as the data transport mechanism in a peer-to-peer backup system.

@@

% TODO: Maybe we should not mention the 1MB test at all.
Our tests for this scenario used 8 virtual machines in SZTAKI Cloud: 1 peer, 5 nodes, 1 tracker, and 1 syslog-ng server. To have a wide range of file sizes, we tested with 1MB, 10MB, 100MB, and 1GB files. In each test instance, the total amount of data backed up was 1GB, so 1000 files, 100 files, 10 files, and 1 file were backed up, respectively. The procedure for each test instance is as follows:

@@

% TODO: This line goes off the right side of the page in the PDF.
\begin{verbatim}
sar -n DEV 1 1 | grep eth0 | grep -v Average | awk '{print $1, $2, $6, $7, "\n";}'
\end{verbatim}

@@

>>>>

\begin{figure}
  \centering
  \includegraphics[scale=0.75]{figures/PeerStoreTrading}
  \caption{Trading protocol of PeerStore \cite{PeerStore} \label{fig:peerStoreTrading}}
\end{figure}

@@

\begin{figure} \label{fig:convergent}
  \centering
  \includegraphics[scale=0.75]{figures/ConvergentEncryption}
  \caption{Convergent Encryption from PeerStore}
\end{figure}

@@

\section{The BitTorrent Protocol} \label{sec:TheBitTorrentProtocol}
BitTorrent is a peer to peer protocol designed for fast file sharing. A peer who wishes to share a file creates a .torrent file containing metadata about the file to be shared. This peer is known as the file's initial \textit{seeder}, or uploader, and through the use of a BitTorrent client, starts \textit{seeding}, or uploading, the file. When a peer acts as a seeder, its sole purpose is to share the file that it is seeding with other peers. Peers who wish to download the file must acquire its .torrent file through an out-of-band channel, which is often a torrent indexer. Peers use the .torrent file to locate the initial seeder through either a centralized tracker or a DHT. These peers then connect to the initial seeder and start downloading the file. These peers are known as \textit{leechers} until they possess a full copy of the file, which is when they, themselves, become seeders. A \textit{swarm} refers to all entities, seeders and leechers, participating in a given torrent.

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.4]{figures/SystemDesign}
  \caption{The design of our system \label{fig:SystemDesign}}
\end{figure}

@@

In order to thoroughly evaluate our system, we identified several variables that could lead to differences in performance. These variables are:

@@

From these variables, we derived a set of test scenarios which allowed us to see how each of them affects the performance of the system. While they are specific to our system, they allow us to evaluate the strengths and weaknesses of BitTorrent Sync (and, by extension, BitTorrent) as the data transport mechanism in a peer-to-peer backup system.

@@

% TODO: Maybe we should not mention the 1MB test at all.
Our tests for this scenario used 8 virtual machines in SZTAKI Cloud: 1 peer, 5 nodes, 1 tracker, and 1 syslog-ng server. To have a wide range of file sizes, we tested with 10MB, 100MB, and 1GB files. In each test instance, the total amount of data backed up was 1GB, so 100 files, 10 files, and 1 file were backed up, respectively. The procedure for each test instance is as follows:

@@

\begin{verbatim}
sar -n DEV 1 1 | grep eth0 | grep -v Average | \
  awk '{print $1, $2, $6, $7, "\n";}'
\end{verbatim}

The \textit{sar} command retrieves the desired network statistics, from which the other commands extract the specific data we are interested in.

@@

In this chapter, we present the reader with the results of the test scenarios outlined in Section \ref{sec:TestScenatios_chap:Methodology} and their analyses.

\section{File Size versus Number of Files} \label{sec:FileSizeversusNumberofFiles_chap:Results}

\subsection{Data Recovery Speed from Peer's Perspective} \label{subsec:DataRecoverySpeedfromPeersPerspective}

Figure \ref{fig:PeerRecoverySpeed} shows the speed of a peer recovering 1 GB of data in three different situations. Each situation differs in how many files comprise the 1 GB payload: one 1 GB file, ten 100 MB files, and one hundred 10 MB files. The moving average of each data series is shown instead of the raw data to reduce visual clutter; the trends are analogous.

\begin{figure}
  \centerline{\includegraphics[scale=.8]{figures/PeerRecoverySpeed}}
  \caption{The speed at which a peer recovered 1 GB worth of data. \label{fig:PeerRecoverySpeed}}
\end{figure}

It is clear that the peer downloaded its data faster when the data was comprised of fewer files. The peer downloaded the fastest when there was only one file to recover, and the slowest when there were one hundred files to recover. This aligns with the theory behind the BitTorrent protocol, which is designed to transfer larger files quicker.

\begin{table}
\begin{center}
    \begin{tabular}{| l | l | l | l |}
    \hline
    & One 1 GB File & Ten 100 MB files & Hundred 10 MB files \\ \hline
    Mean & 8318.52 & 7210.29 & 5405.28 \\ \hline
    Standard Deviation & 3009.26 & 3444.88 & 1738.38 \\ \hline
    Median & 9470.31 & 8599.82 & 5923.23\\ \hline
    Max & 11691.1 & 10926.5 & 8290.26\\ \hline
    \end{tabular}
    \caption{Statistics on Peer Data Recovery Speed (in kB/s)}
    \label{tab:PeerRecoverySpeed}
\end{center}
\end{table}

It is worth noting that the increase in speed does not scale linearly with the size of the file. Although the size of the files increased by the same order of magnitude between the trials, the increase in recovery speed seems to have diminishing returns. Table \ref{tab:PeerRecoverySpeed} provides relevant statistics. The mean recovery speed increased by 33\% between the "Hundred 10 MB Files" trial and the "Ten 100 MB Files" trial. Compare this to an increase of 15\% between the "Ten 100 MB Files" trial and the "One 1 GB File" trial. Furthermore, the total time it took to recover the file between the former trials improves by 60 seconds, whereas the recovery time between the latter trials improves by only 3 seconds.

\subsection{Data Recovery Speed from Nodes' Perspective}

Figure \ref{fig:NodeRecoverySpeed} shows the combined speed at which the nodes are uploading a 1 GB file to a peer. A moving average is used to reduce visual clutter. We can see a significant improvement between the combined upload speed achieved by all nodes compared to the individual upload speeds of each node.

\begin{figure}
  \centerline{\includegraphics[scale=.8]{figures/NodeRecoverySpeed}}
  \caption{The combined speed of nodes uploading 1 GB worth of data to a peer. \label{fig:NodeRecoverySpeed}}
\end{figure}

If the peer was restricted to recovering its data from only one node, its download speed would be limited by the node's upload speed. Table \ref{tab:NodeRecoverySpeed} lists the relevant upload speed statistics of the nodes. At best, the peer could download at an average of 2813 kB/s. However, by using BitTorrent as a data transfer protocol instead of a direct download, the peer's download speed is limited by the \textit{sum} of each node's upload speed: 7981 kB/s. This means, on average, the peer can theoretically download 280\% faster!\footnote{This number is based on the mean. The median suggests that the peer can theoretically download 207\% faster.}

\begin{table}
\begin{center}
    \begin{tabular}{| l | l | l | l | l | l |}
    \hline
    & Node 1 & Node 2 & Node 3 & Node 4 & Node 5 \\ \hline
    Mean & 2813.57 & 1441.16 & 1780.92 & 1417.63 & 530.91 \\ \hline
    Standard Deviation & 2316.67 & 1765.76 & 1758.64 & 1837.77 & 913.80 \\ \hline
    Median & 2281.84 & 703.98 & 1169.43 & 384.91 & 195.90 \\ \hline
    Max & 9000.08 & 7516.42 & 7600.65 & 7071.78 & 5147.33 \\ \hline
    \end{tabular}
    \caption{Statistics on Node Upload Speed (in kB/s)}
    \label{tab:NodeRecoverySpeed}
\end{center}
\end{table}

\subsection{Entities Backing Up a 1 GB File}

Figure \ref{fig:EntityBackupSpeed} shows the speeds at which each participating entity is backing up a 1 GB file.  We use a 20 period moving average to reduce the visual clutter while keeping the overall trends the same.\footnote{The use of a 20 period moving average significantly alters the peaks and valleys of each line. This is acceptable for our purposes because we are only concerned with their existence and relative placements, not their individual details.}

\begin{figure}
  \centerline{\includegraphics[scale=.8]{figures/EntityBackupSpeed}}
  \caption{The speed at which each entity backs up a 1 GB file. \label{fig:EntityBackupSpeed}}
\end{figure}

We confirm that the nodes, in addition to the peer, are actively backing up the peer's file. This is the expected behavior of the BitTorrent protocol, which allows all users in a swarm to share data with each other. BitTorrent Sync, which uses a modified BitTorrent protocol, also seems to leverage this feature. % TODO why is the peer consistently uploading faster than nodes? seeder vs leecher maybe?

We believe the initial ~30 seconds of inactivity is due to BitTorrent Sync indexing the 1 GB file before it can transfer it.

\section{Effects of Churn}

\section{Number of Replicas}
